Searching for long-period transiting planets in the \kepler\ light curves using supervised classification

### 2015-04-08 document/ms.tex
one or two more words
WORDS WORDS WORDS

\newcommand{\figref}[1]{\ref{fig:#1}}
\newcommand{\Fig}[1]{\figurename~\figref{#1}}
\newcommand{\fig}[1]{\Fig{#1}}
\newcommand{\figlabel}[1]{\label{fig:#1}}
\newcommand{\Tab}[1]{Table~\ref{tab:#1}}
\newcommand{\tab}[1]{\Tab{#1}}
\newcommand{\tablabel}[1]{\label{tab:#1}}
\renewcommand{\eqref}[1]{\ref{eq:#1}}
\newcommand{\Eq}[1]{Equation~(\eqref{#1})}
\newcommand{\eq}[1]{\Eq{#1}}
\newcommand{\eqalt}[1]{Equation~\eqref{#1}}
\newcommand{\eqlabel}[1]{\label{eq:#1}}
\newcommand{\sectionname}{Section}
\newcommand{\Sect}[1]{\sectionname~\ref{sect:#1}}
\newcommand{\sect}[1]{\Sect{#1}}
\newcommand{\sectalt}[1]{\ref{sect:#1}}
\newcommand{\App}[1]{Appendix~\ref{sect:#1}}
\newcommand{\app}[1]{\App{#1}}
\newcommand{\sectlabel}[1]{\label{sect:#1}}

@@

We propose a transit search method that shares some similarities with the PH
model but, instead of human classifiers, we use a supervised classification
model trained on large numbers of simulated transit signals injected into
real \kepler\ light curves.
This procedure is motivated because, by definition, most randomly selected
sections of light curve contain no transits and there is an excellent physical
model for the signals of interest.
If we can train a classification model to robustly separate light curve
sections with transits from sections without then we should be able to use
this in-place of the humans to find single transit events.

\section{Estimated yield}

@@

\section{Data preparation}

The \kepler\ Mission measured photometry for about 190,000 stars at half-hour
cadence for a baseline of over four years.
We aim to search these time series for single transits of long-period planets
and single eclipses of binary stars.
These data are made available on
MAST\footnote{\url{https://archive.stsci.edu/kepler/}} and, for each target,
we downloaded the full set of light curve files provided by Data Release 24
\citep{Thompson:2015}.
From these files, we extracted the PDC time series and split them into
``sections'' with no more than two contiguous missing or flagged data points.
The PDC light curves have been corrected for the instrumental effects caused
by the spacecraft using a data-driven model of the focal plane
\citep{Stumpe:2012, Smith:2012}.
Crucially, an attempt is also made by the PDC procedure to remove sharp
instrumental artifacts like ``sudden pixel sensitivity dropouts (SPSDs)''.

@@

Any remaining missing data points are filled in using linear interpolation
resulting in sections of contiguous time series measurements with nearly
uniform time sampling.
This procedure yields about 150 \emph{light curve sections} for the typical
\kepler\ target.

@@

\section{Random forest classification}

A common task in the machine learning literature is called \emph{supervised
classification} where the goal is to separate objects into classes
represented by sets of labeled examples.
One example from the astronomy literature is the problem of star--galaxy
separation (\dfmtodo{cite}).
In this problem, there are a set of images or photometric measurements that
have been classified---by some other method---as either a star or galaxy
observation and the goal is to transfer these labels to a set of observations
that have not yet been classified.
For a more in-depth discussion of the application of these techniques in
astronomy, the interested reader is directed to \citet{Ivezic:2013}.

The Random Forest (RF) classification model \citep{Breiman:2001} has become
very popular in the machine learning community.
It has also been applied with great success in astronomy \citep[for
example]{Richards:2011, Richards:2012} \dfmtodo{others}.
The RF model works by fitting an ensemble of decision trees to randomly
selected subsamples of the training dataset.
Each tree in the forest is ``grown'' by greedily choosing decision boundaries
that optimize the separation of the training data in randomly selected
subsets of the features until the separation is complete, each leaf contains
only one class or, at most, a fixed number of samples.
At each branch, the decision boundary is set by maximizing either the entropy
or the ``gini'' coefficient.

For the problem of single transit search, our problem can be phrased as a
classification problem where we want to assign every $K$ data point long chunk
of light curve into either the \texttt{transit} or \texttt{no transit} class.
This problem doesn't fall under the standard format of a supervised
classification problem because very few long-period transits have actually
been observed or classified.
We do think, however, that we have a good physical generative model for the
signal of interest and most of the observations of each star have no transits.
Therefore, we can train the model on simulated signals.

We use the \project{scikit-learn}\citep{Pedregosa:2011} implementation of the
RF classification algorithm\footnote{Specifically we use the
\texttt{RandomForestClassifier} object;
\url{http://scikit-learn.org/stable/modules/generated/sklearn.ensemble.RandomForestClassifier.html}.}.

\section{Search procedure}\sectlabel{search}

To search for single transits in the light curve of a specific star, we fit
three random forest models.

@@

We need to train the classification model on simulated signals.
To simulate these training and validation examples, we randomly select
sections of 121 contiguous flux measurements and simulate the transit signal
induced by the physical orbit of a large planet.
The specific distribution of simulated parameters is listed in \dfmtodo{some
table}.
In the simulations, we take limb darkening and integration time into account
\dfmtodo{cite}.
Since we expect most of the systematic variability to be time-reversible, we
augment the training set with simulations injected into the reversed time
series \dfmtodo{cite some examples of rotational invariance}.
The training set also includes an equal number of negative examples---time
series chunks of the same length without a simulated transit injection.
By virtue of the fact that we are specifically searching for single transits
it is safe to assume that none of the negative samples from the other light
curve sections will have no transits.

@@

\paragraph{Features}

@@

\paragraph{Train}

\paragraph{Validate}

\paragraph{Test}

@@

\section{Parameter estimation}

@@

\section{Population inference}

@@

\section{Results}

\section{Discussion}

>>>>

\newcommand{\figref}[1]{\ref{fig:#1}}
\newcommand{\Fig}[1]{\figurename~\figref{#1}}
\newcommand{\fig}[1]{\Fig{#1}}
\newcommand{\figlabel}[1]{\label{fig:#1}}

\newcommand{\Tab}[1]{Table~\ref{tab:#1}}
\newcommand{\tab}[1]{\Tab{#1}}
\newcommand{\tablabel}[1]{\label{tab:#1}}

\renewcommand{\eqref}[1]{\ref{eq:#1}}
\newcommand{\Eq}[1]{Equation~(\eqref{#1})}
\newcommand{\eq}[1]{\Eq{#1}}
\newcommand{\eqalt}[1]{Equation~\eqref{#1}}
\newcommand{\eqlabel}[1]{\label{eq:#1}}

\newcommand{\sectionname}{Section}
\newcommand{\sectref}[1]{\ref{sect:#1}}
\newcommand{\Sect}[1]{\sectionname~\sectref{#1}}
\newcommand{\sect}[1]{\Sect{#1}}
\newcommand{\sectalt}[1]{\sectref{#1}}
\newcommand{\App}[1]{Appendix~\sectref{#1}}
\newcommand{\app}[1]{\App{#1}}
\newcommand{\sectlabel}[1]{\label{sect:#1}}

@@

We propose a transit search method that shares some similarities with the PH
model but, instead of human classifiers, we use a supervised classification
model trained on large numbers of simulated transit signals injected into
real \kepler\ light curves.
This procedure is motivated because, by definition, most randomly selected
sections of light curve contain no transits and there is an excellent physical
model for the signals of interest.
If we can train a classification model to robustly separate light curve
sections with transits from sections without then we should be able to use
this in-place of the humans to find single transit events.
This method is novel as a means to transit search and it differs substantially
from the standard methods.

Automated classification processes have been hugely successful in the machine
learning literature when applied to the right datasets \dfmtodo{cite some
things}.
Many astronomical problems can't be naturally solved by the standard machine
learning models because of their heterogeneous character and heteroscedastic
uncertainties.
The search for single transits, however, is an excellent use case when applied
with care.
The noise in the light curve of a given star is relatively consistent for the
full baseline of observation and we have a very good physical generative model
for the transit signal so we can simulate arbitrarily large sets of training
examples.

In the training step, a Random Forest (RF) classifier is trained to
distinguish---based only on the light curve itself---between sections of light
curve that contain transits from sections of the same size that do not.
This model is then used to find candidates in left-out sections of the light
curve.
In principle, this model can capture arbitrarily complicated non-linear
structure in the decision space and if the training set is large and complete
it will ``learn'' the shape of a physical transit.
In practice, classification algorithms like this also require a complete set
of ``negative'' training examples so further vetting using physical models
is necessary.

In \sect{est}, we extrapolate a recent model of the distribution of exoplanets
to the long periods considered in this \paper.
Using an approximate but realistic model for the detection efficiency of a
search for single transits, we estimate that $\sim 60$ single transit events
should be detectable in the archival \kepler\ dataset.
In \sect{data}, we describe the data and outline the structure exploited by
our transit search procedure.
In \sect{rfc}, we briefly describe the Random Forest classification framework
and in \sect{search} we outline the steps used to apply this model for transit
search.
This method requires the setting of many tuning parameters and the
optimization of these choices is both computationally and intellectually
challenging so in \sect{tuning} we discuss some steps in this direction.
In \sect{demo}, we demonstrate the feasibility of this method, present a
single transit discovery, and indicate some interesting weaknesses of the
method.
In \sectionname s~\sectref{params} and~\sectref{exopop}, we describe the
parameter estimation that is possible given a single transit and outline the
prospects for exoplanet population inference given a catalog of single transit
discoveries.

\section{Estimated yield}\sectlabel{est}

@@

\section{Data preparation}\sectlabel{data}

The \kepler\ Mission measured photometry for about 190,000 stars at half-hour
cadence for a baseline of over four years.
We aim to search these time series for single transits of long-period planets
and single eclipses of binary stars.
These data are made available on
MAST\footnote{\url{https://archive.stsci.edu/kepler/}} and, for each target,
we downloaded the full set of long cadence light curve files provided by Data
Release 24 \citep{Thompson:2015}.
From these files, we extracted the PDC time series and split them into
``sections'' with no more than ten contiguous missing or flagged data points.
The PDC light curves have been corrected for the instrumental effects caused
by the spacecraft using a data-driven model of the focal plane
\citep{Stumpe:2012, Smith:2012}.
Crucially, an attempt is also made by the PDC procedure to remove sharp
instrumental artifacts like ``sudden pixel sensitivity dropouts (SPSDs)''.
The success rate of this correction procedure is much higher than in earlier
data releases but, as discussed in \sect{demo}, there remain some cases that
are not properly accounted for.

@@

Any remaining missing data points are filled in using linear interpolation
resulting in sections of contiguous time series measurements with nearly
uniform time sampling.
This procedure yields about 100 \emph{light curve sections} for the typical
\kepler\ target.

@@

\section{Random forest classification}\sectlabel{rfc}

A common task in the machine learning literature is called \emph{supervised
classification} where the goal is to separate objects into classes
represented by sets of labeled examples.
In the astronomy literature, supervised classification has been used for
variable star classification \citep{Richards:2011} star--galaxy separation
\citep{Fadely:2012}, galaxy morphology prediction \citep{Dieleman:2015}, and
other applications.
In each of these problems, there are a set of measurements that have been
assigned classes---by some other method; often manually---and the goal is to
transfer these labels to a set of observations that have not yet been
labeled.
For a more in-depth discussion of the application of these techniques in
astronomy, the interested reader is directed to \citet{Ivezic:2013}.

In our problem of transit search, we want to ``label'' sections of light curve
with either the \texttt{transit} or \texttt{no transit} class.
This problem doesn't fall under the standard format of a supervised
classification problem because very few long-period transits have actually
been observed or classified.
We do think, however, that we have a good physical generative model for the
signal of interest and most of the observations of each star have no transits.
Therefore, we train the model on simulated signals.

The Random Forest (RF) classification model \citep{Breiman:2001} is a popular
model in the machine learning community where it is a common go-to model for
basic classification tasks.
It has also been applied with great success in astronomy \citep[for
example]{Richards:2011, Richards:2012} \dfmtodo{others}.
The RF model works by fitting an ensemble of decision trees to randomly
selected subsamples of the training dataset.
Each tree in the forest is ``grown'' by greedily choosing decision boundaries
that optimize the separation of the training data in randomly selected
subsets of the features until the separation is complete, each leaf contains
only one class or, at most, a fixed number of samples.
At each branch, the decision boundary is set by maximizing either the entropy
or the ``gini'' coefficient.

We use the \project{scikit-learn} \citep{Pedregosa:2011} version of the RF
classification algorithm\footnote{Specifically we use the
\texttt{RandomForestClassifier} object;
\url{http://scikit-learn.org/stable/modules/generated/sklearn.ensemble.RandomForestClassifier.html}.}
implemented in \project{Python}.
This implementation has state-of-the-art speed and
performance\footnote{\url{http://blog.explainmydata.com/2014/03/big-speedup-for-random-forest-learning.html}}
while maintaining a flexible and user-friendly interface.

\section{Search methodology}\sectlabel{search}

Now, we will apply the RF classification model to search for single transits
in \kepler\ light curves.
The basic structure of the problem, applied to the light curve of a single
star, is:
\begin{enumerate}

{\item {\bf train} a classifier on simulated signals injected into a subset of
the light curves sections,}

{\item {\bf validate} the model on signals injected into a different subset of
the data to estimate the precision of the results,}

{\item {\bf test} the classifier by applying it to a final subset of the
data---disjoint from the previous two sets---to predict the class
(\texttt{transit} or \texttt{no transit}) of each light curve section, and}

{\item iterate until the entire light curve has been classified.}

\end{enumerate}
This means that searching a single light curve involves training, validating,
and testing at least 3 RF models.
In practice, we find that it is sufficient to split the data (intelligently)
into 3 disjoint subsets so that we need only train 3 classifiers for each
star.

A key assumption of this model is that the classes are sufficiently
represented in feature space.
In other words, the training set must span both of the classes.
Given a large enough training set, this is not difficult for the
\texttt{transit} class; we can simulate a transit for any combination of
physical parameters.
This is much more difficult for the \texttt{no transit} class.
Since we don't have a generative model for the stellar and instrumental
variability or other artifacts, we cannot simulate negative examples that
cover the full space of possibilities.
Instead we are using sections of real light curves as the \texttt{no transit}
examples.
This means that there are a finite number of samples available for training
and validation, and if there is a qualitatively different signal in the test
set that is not generated by a transit, the model is at risk of unpredictably
misclassifying that section.
This problem can be largely mitigated by careful splitting of the dataset---as
described below---but there are inevitably some false signals that are not
properly accounted for.

@@

We need to train the classification model on simulated signals.
To simulate these training and validation examples, we randomly select
sections of 121 contiguous flux measurements and simulate the transit signal
induced by the physical orbit of a large planet.
The specific distribution of simulated parameters is listed in \dfmtodo{some
table}.
In the simulations, we take limb darkening and integration time into account
\dfmtodo{cite}.
Since we expect most of the systematic variability to be time-reversible, we
augment the training set with simulations injected into the reversed time
series \citep[for example][]{Dieleman:2015}.
The training set also includes an equal number of negative examples---time
series chunks of the same length without a simulated transit injection.
By virtue of the fact that we are specifically searching for single transits
it is safe to assume that none of the negative samples from the other light
curve sections will have no transits.

@@

\paragraph{Feature selection}

@@

@@

\section{Parameter estimation}\sectlabel{params}

@@

\section{Prospects for population inference}\sectlabel{exopop}

@@

\section{Discussion}\sectlabel{discussion}
